### 2014-11-03 evaluation.tex
Major addition to implementation

\item \textbf{LWW register}: A last-write-wins register.

@@

\item \textbf{Online store}: Models an online store with shopping cart and
dynamically changing item prices. Checkout process verifies that the custormer
only pays the accepted price. \item \textbf{RuBis}: An ebay-like auction
site~\cite{}.

\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
Twissandra~\cite{}.

@@

\subsection{Contract Characteristics}

@@

The distribution of contracts in these applications is given in
Table~\ref{tab:ctrts}. We see that majority of the operations are classified as
eventually consistent. Typically, causally consistent operations tend to be
reads that require to see operations in the same session or causally preceding
effects. Transactions are useful for maintaining cross-object consistency and
integrity constraints. For example, we maintain foregin key relationship
between tables using RC isolation level on the insertion transaction, and MAV
on the read transaction; if a MAV transaction witnesses the foreign key field
on table, then it \emph{will} witness the corresponding primiary key row in the
referenced table.

RR transaction comes in handy for maintaining the integrity of follow user
functionality in microblog. If user $A$ is followed by $B$, then we add a
\emph{followedBy} edge from $A$ to $B$ and a \emph{following} edge in the other
direction. The integrity constraint is that either both the edges are added, or
none. RR transaction isolation level is used to ensure correctness when
concurrently $A$ blocks $B$, and $B$ follows $A$. By operating on a snapshot of
consistent state, the block user transaction will not be interfered by the
concurrent follow user transaction.

The proof obligations associated with contract classfication is discarged
through the Z3 SMT Solver. In particular, the contract classification process
is completely performed compile time and has no overheads at runtime. Across
our benchmarks, classifying a contract took 11.5 milliseconds on average.

@@

\subsubsection{Configuration}

We deploy \name applications in \emph{clusters} where each cluster is composed
of 5 fully replicated Cassandra replicas within the same datacenter. We
instantiate one shim layer node for every cassandra replica, an place it on the
same VM as the cassandra replica. The clients are instantiated on the same data
center as the store, and run the transactions. We deploy the each node in the
cluster cluster on \cf{c3.4xlarge} Amazon EC2 instances with 16 virtual CPUs,
30GB memory, and 320GB of SSD storage. Our shim layer nodes are multi-threaded,
and we allocate 8 CPUs for each shim layer node. The clients also run on
\cf{c2.4xlarge} instances. We call this deployment \cf{1DC}. For our WAN
experiments, we instantiate 2 clusters (each with 5 nodes), one in US-east (N.
Virginia) and another in US-west (Oregon). The average latency between the
regions was 85 ms. We call this deployment \cf{2DC}.

\subsubsection{LWW Register}

\begin{figure*}
  \centering
  \subfigure[Operation performance 1DC]{\label{grf:LWW-perf-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC.pdf}}
  \subfigure[Operation performance 2DC]{\label{grf:LWW-perf-2DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-2DC.pdf}}
  \subfigure[Transaction performance 1DC]{\label{grf:LWW-txn-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC-txn.pdf}}
  \subfigure[Mutations in transactions]{\label{grf:LWW-txn-wp}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn-wp-vs-tp.pdf}}
	\caption{Performance of LWW Register RDT.}
  \label{grf:LWW_perf}
\end{figure*}

We analyze the performance of different consistency classifications in \name
using the LWW register. Our client workload was generated using YCSB
benchmark~\cite{}, with the default workload of 50\% reads and 50\% writes,
with keys uniformly chosen from a set of 100,000 keys. We varied the number of
clients from 128 to 1024, with each workload running for 180 seconds.
Last-writer-wins is the conflict resolution policy used by Cassandra. Hence, we
are able to directly compare the performance of \name implementation against
Cassandra.

Figure~\ref{grf:LWW-perf-1DC} shows the throughput vs. latency graph for the
different consistency levels in \cf{1DC} configuration. We also include
Cassandra, which provides basic eventual consistency. Recall that \name's
eventual consistency is stronger than the guarantees provided by Cassandra
(Section~\ref{sec:store_sem}). With 512 clients, the average latency of
operations on Cassandra was 32.24 milliseconds. EC and CC under \name was 30\%
and 57\% slower whereas strong consistency was 241\% slower. EC(SC) was able to
achieve 80\%(40\%) of throughput of vanilla Cassandra. The results show that
while \name provides an expressive programming model with stronger consistency
guarantees, its performance is comparable to vanilla Cassandra. However, SC has
a huge performance hit due to coordination between all the replicas for
obtaining object leases as well as reading and writing to all replicas. This
overhead worsens in a WAN setting. Figure~\ref{grf:LWW-perf-2DC} shows the
throughput vs. latency graph under 2DC configuration. While the Cassandra, EC
and CC numbers largely remain the same when compared to 1DC configuration, SC
is 6$\times$ to 12$\times$ higher latency and lower throughput than than the
1DC case.

We compare the performance of different transaction isolation level choices in
Figure~\ref{grf:LWW-txn-1DC}. The numbers were obtained under 1DC
configuration. The YCSB workload was modified to issue 10 operations per
transaction. Each operation is assumed to have eventual consistency. \cf{NoTxn}
corresponds to a configuration that does not use transactions. Compared to this
RC is only 12\% shower in terms of latency with 512 clients, where as RR is
2.3X slower. The difference between RC and \cf{NoTxn} is due to the meta-data
overhead of recording transaction information in the object state. For RR
transaction, The cost of capturing and maintaining the snapshot in an RR
transaction is the biggest source of overhead.

Moreover, the size of the snapshot is proportional to the the number of
different objects accessed by the session as well as the number of updates.
Hence, with increasing proportion of writes, the overhead of RR transaction
increases as shown in Figure~\ref{grf:LWW-txn-wp}. The numbers were obtained
with 128 clients. Compared to an RC transaction whose throughput decreasess
45\% as the write proportion increases from 0\% to 50\%, RR transaction's
throughput drops 78\% in the same interval.

\subsection{Bank Account}

\begin{figure}
  \centering
	\includegraphics[width=0.5\columnwidth]{Figures/BA-legend.pdf}
  \subfigure[Latency 1DC]{\label{grf:BA-lat}\includegraphics[width=0.24\textwidth]{graphs/BA-lat.pdf}}
  \subfigure[Throughput 1DC]{\label{grf:BA-tp}\includegraphics[width=0.23\textwidth]{graphs/BA-tp.pdf}}
	\caption{Performance of Bank Account RDT.}
  \label{grf:BA_perf}
\end{figure}

Figure~\ref{grf:BA-lat} and~\ref{grf:BA-tp} analyzes the latency and throughput
of operations in bank account example as we increase the number of clients in
\cf{1DC} configuration. The benchmark uniformly chose from 100,000 keys, where
the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance. Here,
the lines marked EC and CC correspond to all operations being assigned EC and
CC consistency levels. These levels compromise correctness as withdraw has to
be and SC operation. The line SC corresponds to a configuration where all
operations are strongly consistent; this ensures application correctness. \name
corresponds to our implementation, which classifies operations based on their
contracts. At 512 clients, \name implementation was within 30\% of latency and
12\% of throughput of EC, while providing correctness, whereas SC operations
had 162\% higher latency and 52\% lower throughput than EC operations. This
illustrates the benefit of fine-grained contract classification in \name.

>>>>

\item \textbf{LWW register}: A last-write-wins register that provides read and
write operations. Each write is associated with a timestamp, which is used to
resolve conflicting concurrent writes -- newer write wins.

@@

\item \textbf{Online store}: Models an online store with shopping cart and
dynamically changing item prices. Checkout process verifies that the custormer
only pays the accepted price.

\item \textbf{RuBis}: An ebay-like auction site~\cite{}. The application allows
users to browse items, bid for items on sale, and pay for items from a wallet
modelled after a bank account.

\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
Twissandra~\cite{}. The application allows adding a new user, adding and
replying to tweets, following, unfollowing and blocking users, and fetching a
user's timeline, userline, followers and following.

@@

@@

The distribution of contracts in these applications is given in
Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
are classified as eventually consistent and RC, respectively. Operation
contracts are used to enforce integrity and visibility constraints on
individual fields in the tables. Transactions are mainly used to consistently
modify and access related fields across tables. The proof obligations
associated with contract classfication is discarged through the Z3 SMT Solver.
In \name, the contract classification process is completely performed compile
time and has no overheads at runtime. Across our benchmarks, classifying a
contract took 11.5 milliseconds on average.

@@

We deploy \name applications in \emph{clusters} where each cluster is composed
of 5 fully replicated Cassandra replicas within the same datacenter. We
instantiate one shim layer node for every cassandra replica, an place it on the
same VM as the cassandra replica. The clients are instantiated on the same data
center as the store, and run the transactions. We deploy the each node in the
cluster cluster on \cf{c3.4xlarge} Amazon EC2 instances with 16 virtual CPUs,
30GB memory, and 320GB of SSD storage. Our shim layer nodes are multi-threaded,
and we allocate 8 CPUs for each shim layer node. The clients also run on
\cf{c2.4xlarge} instances. We call this \cf{1DC} configuration. For our
geo-distributed experiments (\cf{2DC}), we instantiate 2 clusters, each with 5
nodes, and place the clusters on US-east (Virginia) and US-west (Oregon). The
average inter-region latency was 85ms.

\begin{figure*}
  \centering
  \subfigure[Bank account latency]{\label{grf:BA-lat}\includegraphics[width=0.252\textwidth]{graphs/BA-lat.pdf}}
  \subfigure[Bank account throughput]{\label{grf:BA-tp}\includegraphics[width=0.24\textwidth]{graphs/BA-tp.pdf}}
  \subfigure[LWW transactions]{\label{grf:LWW-txn}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn.pdf}}
  \subfigure[Impact of summarization]{\label{grf:summarization}\includegraphics[width=0.235\textwidth]{graphs/summarization.pdf}}
	\caption{Quelea Performance.}
  \label{grf:LWW_perf}
\end{figure*}

Figure~\ref{grf:BA-lat} and~\ref{grf:BA-tp} analyzes the latency and throughput
of operations in bank account example as we increase the number of clients in
\cf{1DC} configuration. Our client workload was generated using YCSB
benchmark~\cite{}. The benchmark uniformly chose from 100,000 keys, where the
operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
corresponds to the default 50:50 read:write mix in YCSB. Here, the lines marked
EC and CC correspond to all operations being assigned EC and CC consistency
levels. These levels compromise correctness as withdraw has to be and SC
operation. The line SC corresponds to a configuration where all operations are
strongly consistent; this ensures application correctness. \name corresponds to
our implementation, which classifies operations based on their contracts.

Both \name and SC ensure correctness. However, with 512 clients, \name
implementation was within 30\% of latency and 12\% of throughput of EC, whereas
SC operations had 162\% higher latency and 52\% lower throughput than EC
operations. In \cf{2DC} configuration (not-shown), the average latency of SC
operations with 512 clients increased by 9.4$\times$ due to the cost of
geo-distributed coordination, whereas \name operations were only 2.2$\times$
slower, mainly due to the increased cost of \cf{withdraw} operations.
Importantly, the latency of \cf{getBalance} and \cf{deposit} remained almost
the same. This illustrates the benefit of fine-grained contract classification
in \name.

We compare the performance of different transaction isolation level choices in
Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
under 1DC configuration. The YCSB workload was modified to issue 10 operations
per transaction, with the default 50:50 read:write mix. Each operation is
assumed to have eventual consistency. \cf{NoTxn} corresponds to a configuration
that does not use transactions. Compared to this RC is only 12\% shower in
terms of latency with 512 clients, where as RR is 2.3X slower. The difference
between RC and \cf{NoTxn} is due to the meta-data overhead of recording
transaction information in the object state. For RR transaction, the cost of
capturing and maintaining the snapshot in an RR transaction is the biggest
source of overhead.

We also compared the performance of EC LWW operations directly against
Cassandra (our backing store), which uses last-writer-wins as the only
convergence semantics. While Cassandra provides no stronger-than-eventual
consistency properties, \name was within 30\%(20\%) of latency(throughput) of
Cassandra with 512 clients, illustrating that the programmers only have to pay
a minimal overhead for the expressive and stronger \name programming model.

Finally, we study the impact of summarization in
Figure~\ref{grf:summarization}. We utilize 128 clients and a single \name
replica, with all the clients operating on the \emph{same} LWW register to
stress test the summarization mechanism. We utilize 50:50 read:write mix, and
33:33:33 EC:CC:SC mix. The shim layer cache (mem) of operations is summarized
every 64 updates, while the updates in the backing store (disk) are summarized
every 4096 updates. Each point in the graph represents the average latency of
the previous 1000 operations. Each experiment is run for 60s.

The results show that without summarization, the average latency of operations
increase exponentially to almost 1 second, and only 13K operations were
performed in a minute. Since every operation has to reduce over the set of all
previous operations, with a ever growing set, the operations take increasingly
more time to complete. With summarization only in memory, the performance still
degrades due to the cost of fetching all previous updates from the backing
store into the shim layer. Fetching the latest updates is essential for SC
operations. With both summarizations enabled, we see that the latency does not
increase over time, and we were able to perform 67K operations. This graph
illustrates the importance and effectiveness of summarization in \name.
